Dark matter halo shape in the Auriga simulations: radial dependence, time evolution and baryon disk alignment

### 2019-07-18 Document/paper.tex
new plots
remove level 3 results

\begin{figure*}
\begin{center}
\includegraphics[width=0.35\textwidth]{Z_Triax_level3_set_A_DM.pdf}
\includegraphics[width=0.35\textwidth]{Z_Triax_level3_set_A_MHD.pdf}
\end{center}
\caption{Axial ratios as a function of time (colour circles) and
  radius (dashed lines). 
  The left/right columns correspond to DMO/MHD simulations, respectively. 
  The colour indicates the redshift at which the shape was measured
  (always at $R_{200}$ at that time.)
  The radial shapes goes from $R_{200}$ down to $\sim 1$ kpc.
  These three halos correspond to a first half the sample for the \texttt{Level3}
  haloes. 
}
\label{fig:triaxial_history_A}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.35\textwidth]{Z_Triax_level3_set_B_DM.pdf}
\includegraphics[width=0.35\textwidth]{Z_Triax_level3_set_B_MHD.pdf}
\end{center}
\caption{Same as Figure \ref{fig:triaxial_history_A}.
  These three halos correspond the second half of the sample for the
  \texttt{Level3} haloes. 
}
\label{fig:triaxial_history_B}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{cumulative_alignment_DM.pdf}
\includegraphics[width=1.0\textwidth]{cumulative_alignment_MHD.pdf}
\end{center}
\caption{Cumulative distribution for the alignments between the
  principal axis of the halo shape and the principal axis of the
  stellar disk in the MHD simulations for radii $\leq 0.5R_{200}$.
  Each panel compares the alingment of the corresponding
  major/middle/minor axis both in the halo and the stellar disk.
  In the upper row the haloes come from the DMO simulation, 
  here the central message is that the shells are aligned to each
  other at every radii.
  In the lower row the haloes come from the MHD simulation and is a
  self-consistent comparison with the disks. 
  In this case the dark matter shells twist and  the degree of the
  halo-disk alignment changes with distance.
  The worst halo-disk alignment happens closer to the disk ($14\pm1$ kpc).
  Interestingly an almost perfect halo-disk alignment happens at an intermediate
  radius of $0.25R_{200}$ ($56\pm 4$kpc); while at a larger
  distance $0.5R_{200}$ the two start to missalign.
}
\label{fig:cumulative_alignment}
\end{figure*}

@@

@@

\subsection{Shape evolution with cosmic time}

In the previous section we stablished that DMO halos at $z=0$ have
a steady and monotonous trend towards decreasing triaxility with
increasing radius.
\cite{VeraCiro11} found that this  radial trend at $z=0$ 
correlates with the halo shape as a function of cosmic time, i.e.
the halo becomes less triaxial as cosmic time progresses. 
The hypotesis put forward by \cite{VeraCiro11} is that the correlation
raises as each radial shell retains memory of the accreting conditions
at the virial radius at the time of collapse.

Here we use the six \texttt{level3} halos to measure to what extent
the presence of baryons affect this memory effect.
In Figures \ref{fig:triaxial_history_A} and
\ref{fig:triaxial_history_B} we follow the shape evolution measured at
$R_{200}$ since redshift $z=1$ to the present, these results are
represened by the coloroured dots.
We compare this temporal evolution against the radial evolution at
$z=0$ represented by the dash black lines.

The left column of Figure \ref{fig:redshift_triaxial} shows the six
halos in the triaxility plane. 
The colour dots show the time evolution while the black line
corresponds to the radial trend at $z=0$.
The left/right columns correspond to the DMO/MHD simulations, respectively.

We start by analyzing to what extent the memory effect is
already present in the DMO sample. 
The first feature that jumps out is that \texttt{Au-21} in the DMO
simulation does not have a simple shape evolution history.
By visual inspection we find that an extended merging substructure
intereacted in the main halo showing up in the shape measurements,
making it more difficult to interpret the memory hypothesis.
For this reason we exclude this halo from the analysis.

For the other five halos we find simpler temporal and spatial
evolution histories. 
Halo \texttt{Au-6} (Figure
\ref{fig:triaxial_history_A}) shows a broad
correlation between the two tracks; close to the cases of halos
\texttt{Aq-4-C} and \texttt{Aq-4-D} in \cite{VeraCiro11}.
The four other halos (\texttt{Au-16} in Figure
\ref{fig:triaxial_history_A}; \texttt{Au-23}, \texttt{Au-24},
\texttt{Au-27} in Figure \ref{fig:triaxial_history_B}) are four
cases where the temporal evolution if followed very closely by the
radial trends.
We conclude that in the DMO sample 4 out of the 5 high resolution
halos closely follow the memory hypothesis trends and the remaning
halo shows a reasonable correlation.

The situation in the MHD simulations is more diverse.
As expected by the results in the previous section the broad location
in the triaxility plane of the temporal and radial evolution in the
MHD simulation is completeley different from its DMO counterpart. 
In terms of the correlations expected by the memory hypothesis we see
two different trends.
We have the case of Halos \texttt{Au-6}, \texttt{Au-16} and
\texttt{Au-27} there is still a reasonable correlation between
temporal and radial evolution. 
However for halos \texttt{Au-23} and \texttt{Au-24} the radial trends
go in directions that are not even covered by the temporal evolution.

Taken at face value, these results show that it is not possible to
generalize the memory hypotesis to dark matter halos in MHD
simulations.
This means that some halos present a amnesia to its formation history,
more specifically there are cases where the radial shape profile at
present time does not have any correlation to its evolutionary
history.

@@

>>>>

%\begin{figure*}
%\begin{center}
%\includegraphics[width=0.35\textwidth]{Z_Triax_level3_set_A_DM.pdf}
%\includegraphics[width=0.35\textwidth]{Z_Triax_level3_set_A_MHD.pdf}
%\end{center}
%\caption{Axial ratios as a function of time (colour circles) and
%  radius (dashed lines). 
%  The left/right columns correspond to DMO/MHD simulations, respectively. 
%  The colour indicates the redshift at which the shape was measured
%  (always at $R_{200}$ at that time.)
%  The radial shapes goes from $R_{200}$ down to $\sim 1$ kpc.
%  These three halos correspond to a first half the sample for the \texttt{Level3}
%  haloes. 
%}
%\label{fig:triaxial_history_A}
%\end{figure*}

%\begin{figure*}
%\begin{center}
%\includegraphics[width=0.35\textwidth]{Z_Triax_level3_set_B_DM.pdf}
%\includegraphics[width=0.35\textwidth]{Z_Triax_level3_set_B_MHD.pdf}
%\end{center}
%\caption{Same as Figure \ref{fig:triaxial_history_A}.
%  These three halos correspond the second half of the sample for the
%  \texttt{Level3} haloes. 
%}
%\label{fig:triaxial_history_B}
%\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{angles_alignment_DM.pdf}
\includegraphics[width=1.0\textwidth]{angles_alignment_MHD.pdf}
\end{center}
\caption{Angles between the
  principal axis of the halo shape and the principal axis of the
  stellar disk in the MHD simulations at four different radii $\leq 0.5R_{200}$.
  Thin lines correspond to each one of the thirty halos in our sample,
  while the thick line traces the median value.
  Each panel compares the alignment of the corresponding
  major/middle/minor axis both in the halo and the stellar disk.
  In the upper row the haloes come from the DMO simulation, 
  showing that the ellipsoids describing the
  shape are constant as a function of radius for the most part of the
  sample.
  In the lower row the haloes come from the MHD simulation providing a 
  self-consistent comparison with the stellar disks. 
  In this case the dark matter shells twist in all the halos.
  The degree of this twisting is different in each halo.
  Interestingly, an almost perfect halo-disk alignment happens across
  the sample at an intermediate radius of $0.25R_{200}$ ($56\pm 4$kpc).
}
\label{fig:cumulative_alignment}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{correlations_angles_alignment_MHD.pdf}
\end{center}
\caption{Changse of angle alignment of the dark
  matter halo and stellar disk at two different radii ($R_{200}/16$
  and $0.25R_{200}$) as a function of the baryonic disk properties
  already explored in Figure \ref{fig:disk_correlations}.  
  Figure \ref{fig:cumulative_alignment} showed that maximum alignment
  occurs at $0.25R_{200}$ while in the inner regions ($R_{200}/16$)
  considerable missalignments occurr.
 at different radii
 and baryonic disk properties. 
 The label with the $\rho$ value corresponds to the Spearman’s rank correlation coefficient. }
\label{fig:alignment_correlations}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{correlations_angles_triaxiality_MHD.pdf}
\end{center}
\caption{Changse of angle alignment of the dark
  matter halo and stellar disk at two different radii ($R_{200}/16$
  and $0.25R_{200}$) as a function of the baryonic disk properties
  already explored in Figure \ref{fig:disk_correlations}.  
  Figure \ref{fig:cumulative_alignment} showed that maximum alignment
  occurs at $0.25R_{200}$ while in the inner regions ($R_{200}/16$)
  considerable missalignments occurr.
 at different radii
 and baryonic disk properties. 
 The label with the $\rho$ value corresponds to the Spearman’s rank correlation coefficient. }
\label{fig:alignment_correlations}
\end{figure*}

@@

%\subsection{Shape evolution with cosmic time}

%In the previous section we stablished that DMO halos at $z=0$ have
%a steady and monotonous trend towards decreasing triaxility with
%increasing radius.
%\cite{VeraCiro11} found that this  radial trend at $z=0$ 
%correlates with the halo shape as a function of cosmic time, i.e.
%the halo becomes less triaxial as cosmic time progresses. 
%The hypotesis put forward by \cite{VeraCiro11} is that the correlation
%raises as each radial shell retains memory of the accreting conditions
%at the virial radius at the time of collapse.

%Here we use the six \texttt{level3} halos to measure to what extent
%the presence of baryons affect this memory effect.
%In Figures \ref{fig:triaxial_history_A} and
%\ref{fig:triaxial_history_B} we follow the shape evolution measured at
%$R_{200}$ since redshift $z=1$ to the present, these results are
%represened by the coloroured dots.
%We compare this temporal evolution against the radial evolution at
%$z=0$ represented by the dash black lines.

%The left column of Figure \ref{fig:redshift_triaxial} shows the six
%halos in the triaxility plane. 
%The colour dots show the time evolution while the black line
%corresponds to the radial trend at $z=0$.
%The left/right columns correspond to the DMO/MHD simulations, respectively.

%We start by analyzing to what extent the memory effect is
%already present in the DMO sample. 
%The first feature that jumps out is that \texttt{Au-21} in the DMO
%simulation does not have a simple shape evolution history.
%By visual inspection we find that an extended merging substructure
%intereacted in the main halo showing up in the shape measurements,
%making it more difficult to interpret the memory hypothesis.
%For this reason we exclude this halo from the analysis.

%For the other five halos we find simpler temporal and spatial
%evolution histories. 
%Halo \texttt{Au-6} (Figure
%\ref{fig:triaxial_history_A}) shows a broad
%correlation between the two tracks; close to the cases of halos
%\texttt{Aq-4-C} and \texttt{Aq-4-D} in \cite{VeraCiro11}.
%The four other halos (\texttt{Au-16} in Figure
%\ref{fig:triaxial_history_A}; \texttt{Au-23}, \texttt{Au-24},
%\texttt{Au-27} in Figure \ref{fig:triaxial_history_B}) are four
%cases where the temporal evolution if followed very closely by the
%radial trends.
%We conclude that in the DMO sample 4 out of the 5 high resolution
%halos closely follow the memory hypothesis trends and the remaning
%halo shows a reasonable correlation.

%The situation in the MHD simulations is more diverse.
%As expected by the results in the previous section the broad location
%in the triaxility plane of the temporal and radial evolution in the
%MHD simulation is completeley different from its DMO counterpart. 
%In terms of the correlations expected by the memory hypothesis we see
%two different trends.
%We have the case of Halos \texttt{Au-6}, \texttt{Au-16} and
%\texttt{Au-27} there is still a reasonable correlation between
%temporal and radial evolution. 
%However for halos \texttt{Au-23} and \texttt{Au-24} the radial trends
%go in directions that are not even covered by the temporal evolution.

%Taken at face value, these results show that it is not possible to
%generalize the memory hypotesis to dark matter halos in MHD
%simulations.
%This means that some halos present a amnesia to its formation history,
%more specifically there are cases where the radial shape profile at
%present time does not have any correlation to its evolutionary
%history.

@@

{{\bf alinenaciones y twisting as a function of environment, 
the difference between accretting and stalled halos, }

%Porciani: stalled halos keep their dark matter density profile
%while accreting halos have. Shells don't pile up in accreting halos, 
%they approximately do in stalled halos}

%Porciani: The stalled halos are the most similar to the MW at least in
%terms of the velocity anisotropy parameter an should be embeded in a
%thick filament of tha large scale structure.

%IDEA PARA YEIMY: medir scatter en stellar to DM halo mass ratio as a
%function of the anisotropy parameter or the persistence parameter in 
%Illustris. Primer proxy para el galaxy formation bias. Esto se puede
%sacar pronto. (ademas otra medida de los compacto que es la distribucion
%estelar)
%esto puede ser comparado con los datos de vipers que tiene un sesgo 
%donde se ven diferentes clustering a misma masa para galaxias mas o
%menos concentradas

%otra idea: calcular la entropia del beta skeleton para distinguir 
%el clustering de dos muestras diferentes. i.e. la entropia para
%caracterizar assembly bias.
